---
author:
  - name: Ben Barnard
    affiliation: Baylor University
    address: >
      Department of Statistical Sciences
      Waco, Tx
    email: ben_barnard@baylor.edu
  - name: Dean Young
    affiliation: Baylor University
title:
  formatted: "\\pkg{rWishart}: An R package for random wishart matrices"
abstract: >
  The abstract of the article.
keywords:
  # at least one keyword must be supplied
  formatted: [keywords, not capitalized, "\\proglang{Java}"]
  plain:     [keywords, not capitalized, Java]
preamble: >
  \usepackage{amsmath}
output: rticles::jss_article
bibliography: [rWishart.bib]
biblio-style: apalike
---
```{r, include = FALSE}
library(rWishart)
```

# Introduction

# Random Wishart Matrices

## Psuedo Wishart Matrices

## Singular Wishart Matrices

## Fractional Wishart Matrices

# Overview and Use of the Package

## Covariance Matrix Class

In this section we discuss a new class of matrix object in R. The primary purpose for this object is not really apparent in the rWishart package and we won't see some of the true advantages to the covariance class object till the covEst package release. When a user calls the rWishart function,

```{r}
S <- rWishart(n = 1, df = 2, Sigma = diag(1, 10), covariance = TRUE, simplify = FALSE)[[1]]
```

they are given the option to generate sample covariance matrices, $S$, where 
$$df * S \sim W(df, \boldsymbol{\Sigma}).$$
In the covariance class matrix we store a new attribute to the matrix object. We store the degrees of freedom as a formula object. The degrees of freedom can be pulled from the covariance matrix class object by using

```{r}
attributes(S)$df
```

## Simulation Study
